Variation in Primary Care Prescribing in Wales due to COVID-19

### 2021-03-04 writing/paper.Rmd
Method outline and all prescribing inital plots

```{r setup, include = FALSE}
## Packages
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(rebus)
library(serCymruTools)
library(papaja)

@@

* Given the presence of seasonal variation and existing trends in levels of prescribing, we used various time series forecasting methods to account for existing trends and seasonal variation when creating the forecasts.

* For each analysis, we created a forecast that aimed to answer the question, "What would the levels of prescribing have been if COVID-19 had not affected Wales?" We then used these forecasts to estimate the changes in levels of prescribing due to COVID-19.

  * We viewed this, somewhat complex, approach as superior to more simple approaches (e.g., carrying last years levels forward, or even using the mean of the last three years).
* We did not assume that the existing processes would be the same for all drugs or all GP practices, therefore, we investigated the fit of several different time series models to the pre-COVID data. 
  * Using Jan 2015 to Feb 2020 data, we fitted several different models and assessed their accuracy using a cross-validated process to reduce the likelihood of overfitting models to the data.
    * Started with 36 months of data, used 6-month horizon as this was the horizon we would be using for the forecasts, 3-month step (to reduce computation time) **Do we want to re-run this with 1-month?**
    * Only interested in one forecast horizon: six-months
    * 
  * Having chosen the "best model" for each practice based on the pre-COVID data, we then used this to forecast the level of prescribing for each practice

* We used time series modelling to forecast Welsh primary care prescribing levels had .
    * We used the `fable` [@R-fable] and the `fable.prophet` [@R-fable.prophet] packages to conduct the time series analyses.
    * Time series linear model
    * Decomposition model
    * Seasonal naïve (with and without drift)
    * Autoregressive integrated moving average [ARIMA; @Box2015]
    * Holt-Winters Additive Model [@Chatfield1978]
    * Prophet [@Taylor2018]
    * Combination models [cf. @Thomson2019]
  * Prescribing quantities log transformed and forecasts use median values to reduce bias that back transformation would introduce when using the mean
  * Model selection
    * RMSE then MAE

@@

all_prescribing_fable_fits <- 
  all_prescribing_fable_forecasts %>% 
  accuracy(all_prescribing_pre_cv19, list(RMSE = RMSE, MAE = MAE, MAPE = MAPE,
                           rmse_skill = skill_score(RMSE),
                           crps_skill = skill_score(CRPS), ACF1 =ACF1,
                           winkler = winkler_score))

all_prescribing_prophet_fits <- 
  all_prescribing_pre_cv19 %>%
  mutate(datename = yearmonth(datename)) %>%
  fill_gaps() %>%
  slice(1:(n()-2), .preserve = TRUE) %>%
  stretch_tsibble(.init = 36, .step = 3) %>%
  prophetModels() %>%
  forecast(h = 6, point_forecast = list(.median = median)) %>%
  accuracy(all_prescribing_pre_cv19, list(RMSE = RMSE, MAE = MAE, MAPE = MAPE,
                           rmse_skill = skill_score(RMSE),
                           crps_skill = skill_score(CRPS), ACF1 =ACF1,
                           winkler = winkler_score))

all_prescribing_combined_fits <- 
  bind_rows(all_prescribing_fable_fits, all_prescribing_prophet_fits) %>%
  arrange(RMSE, MAE)

all_prescribing_best_model <- 
  all_prescribing_combined_fits %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE)

@@

plan(sequential)
```

```{r, all-prescribing-model-accuracy-hist, fig.cap="All prescribing, model accuracy"}
all_prescribing_combined_fits %>% 
  ggplot(aes(x = reorder(.model, RMSE, mean), y = RMSE)) +
  geom_col()
```

@@

```{r, all-prescribing-forecast-plot, fig.cap="All prescribing, best model fit and forecast"}
all_prescribing %>% 
  ggplot(aes(x = datename)) +
  geom_line(aes(y = n, col = "Observed")) +
  geom_line(aes(y = med, col = "Forecast"), 
            data = all_prescribing_covid_forecast) +
  geom_line(aes(y = fitted, col = "Fitted"), 
            data = all_prescribing_best_cv_model %>% 
              group_by(datename) %>% 
              summarise(fitted = mean(.fitted, na.rm = TRUE))) +
  geom_vline(xintercept = ymd("2020-03-01"), lty = 3)
```

>>>>

```{r setup, include = FALSE}
## Packages
devtools::install_github("https://github.com/w-hardy/serCymruTools")
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(fasster)
library(feasts)
library(furrr)
library(rebus)
library(serCymruTools)
library(papaja)

@@

* Given the presence of seasonal variation and existing trends in levels of prescribing, we used various time series forecasting methods to account for existing trends and seasonal variation when creating the forecasts.
* For each analysis, we created a forecast that aimed to answer the question, "What would the levels of prescribing have been if COVID-19 had not affected Wales?" We then used these forecasts to estimate the changes in levels of prescribing due to COVID-19. 
* We viewed this, somewhat complex, approach as superior to more simple approaches (e.g., carrying last years levels forward, or even using the mean of the last three years).

### Forecasting models

* We did not assume that the existing processes would be the same for all drugs or all GP practices, therefore, we investigated the fit of several different time series models to the pre-COVID data. 
* We used the `fable` [@R-fable], `fasster` [@R-fasster], and the `fable.prophet` [@R-fable.prophet] packages to conduct the time series analyses.
  * Time series linear model
  * Decomposition model
  * Seasonal naïve (with and without drift)
  * Autoregressive integrated moving average [ARIMA; @Box2015]
  * Holt-Winters Additive Model [@Chatfield1978]
  * Prophet [@Taylor2018]
  * Combination models [cf. @Thomson2019]
* Prescribing quantities log transformed and forecasts use median values to reduce bias that back transformation would introduce when using the mean
* Using Jan 2015 to Feb 2020 data, we fitted several different models and assessed their accuracy using a cross-validated process to reduce the likelihood of overfitting models to the data.
  * Started with 36 months of data, used 6-month horizon as this was the horizon we would be using for the forecasts, 3-month step (to reduce computation time) **Do we want to re-run this with 1-month?**
    * Only interested in one forecast horizon: six-months

### Model selection

* Ranked by RMSE and Winkler [@Winkler1972], then the best mean ranking model is chosen
  * RMSE is scale dependant. Is suitable for comparing models of the same data.
  * Winkler score penalises models that have observations falling outside the prediction interval, thus rewards models with narrow intervals
* This combination accounts for both the point and distributional accuracy of the forecasts

### Forecasts

* Having chosen the "best model" for each practice based on the pre-COVID data, we then used this model to forecast the level of prescribing for each practice

@@

all_prescribing_fable_fits <- 
  all_prescribing_fable_forecasts %>% 
  accuracy(all_prescribing_pre_cv19, list(RMSE = RMSE, MAE = MAE, MAPE = MAPE,
                           rmse_skill = skill_score(RMSE),
                           crps_skill = skill_score(CRPS), ACF1 = ACF1,
                           winkler = winkler_score))

all_prescribing_prophet_fits <- 
  all_prescribing_pre_cv19 %>%
  mutate(datename = yearmonth(datename)) %>%
  fill_gaps() %>%
  slice(1:(n()-2), .preserve = TRUE) %>%
  stretch_tsibble(.init = 36, .step = 3) %>%
  prophetModels() %>%
  forecast(h = 6, point_forecast = list(.median = median)) %>%
  accuracy(all_prescribing_pre_cv19, list(RMSE = RMSE, MAE = MAE, MAPE = MAPE,
                           rmse_skill = skill_score(RMSE),
                           crps_skill = skill_score(CRPS), ACF1 = ACF1,
                           winkler = winkler_score))

all_prescribing_combined_fits <- 
  bind_rows(all_prescribing_fable_fits, all_prescribing_prophet_fits) %>%
  rankModels()

all_prescribing_best_model <- 
  all_prescribing_combined_fits %>% 
  slice(1, .preserve = TRUE)

@@

plan(sequential)

```{r, all-prescribing-model-accuracy-hist, fig.cap="All prescribing, model accuracy"}
p1 <- all_prescribing_combined_fits %>% 
  na.omit() %>% 
  ggplot(aes(x = reorder(.model, rank, sort), y = RMSE)) +
  geom_col()

p2 <- all_prescribing_combined_fits %>% 
  na.omit() %>% 
  ggplot(aes(x = reorder(.model, rank, sort), y = winkler)) +
  geom_col()

cowplot::plot_grid(p1, p2)
```

@@

```{r, all-prescribing-forecast-plot, fig.cap="All prescribing, best model fit and forecast"}
all_prescribing %>% 
  ggplot(aes(x = datename)) +
  geom_line(aes(y = n, col = "Observed", fill = "Observed")) +
  geom_line(aes(y = med, col = "Forecast", fill = "Forecast"), 
            data = all_prescribing_covid_forecast) +
  geom_ribbon(aes(ymin = lb_80, ymax = ub_80, 
                   fill = "Forecast"), alpha = .5,
              data = all_prescribing_covid_forecast) +
  geom_line(aes(y = fitted, col = "Fitted", fill = "Fitted"), 
            data = all_prescribing_best_cv_model %>% 
              group_by(datename) %>% 
              summarise(fitted = median(.fitted, na.rm = TRUE))) +
  geom_vline(xintercept = ymd("2020-03-01"), lty = 3)
```

### 2021-03-04 writing/paper.tex
Method outline and all prescribing inital plots

  \begin{itemize}
  \tightlist
  \item
    We used the \texttt{fable} (O'Hara-Wild et al., 2021a) and the \texttt{fable.prophet} (O'Hara-Wild, 2020) packages to conduct the time series analyses.
  \item
    Time series linear model
  \item
    Decomposition model
  \item
    Seasonal naïve (with and without drift)
  \item
    Autoregressive integrated moving average (ARIMA; Box et al., 2015)
  \item
    Holt-Winters Additive Model (Chatfield, 1978)
  \item
    Prophet (Taylor \& Letham, 2018)
  \item
    Combination models (cf.~Thomson et al., 2019)
  \item
    Prescribing quantities log transformed and forecasts use median values to reduce bias that back transformation would introduce when using the mean
  \item
    Model selection

    \begin{itemize}
    \tightlist
    \item
      RMSE then MAE
    \end{itemize}
  \end{itemize}
\end{itemize}

@@

We used R (Version 4.0.2; R Core Team, 2020) and the R-packages \emph{dplyr} (Version 1.0.4; Wickham, François, et al., 2021), \emph{fable} (Version 0.3.0; O'Hara-Wild et al., 2021a, 2021b; O'Hara-Wild, 2020), \emph{fable.prophet} (Version 0.1.0; O'Hara-Wild, 2020), \emph{fabletools} (Version 0.3.0; O'Hara-Wild et al., 2021b), \emph{feasts} (Version 0.1.7; O'Hara-Wild et al., 2021c), \emph{forcats} (Version 0.5.1; Wickham, 2021), \emph{furrr} (Version 0.2.2; Vaughan \& Dancho, 2021), \emph{future} (Version 1.21.0; Bengtsson, 2020), \emph{ggplot2} (Version 3.3.3; Wickham, 2016), \emph{lubridate} (Version 1.7.10; Grolemund \& Wickham, 2011), \emph{MplusAutomation} (Version 0.8; Hallquist \& Wiley, 2018), \emph{papaja} (Version 0.1.0.9942; Aust \& Barth, 2020), \emph{prophet} (Version 0.6.1; O'Hara-Wild, 2020; Taylor \& Letham, 2020), \emph{purrr} (Version 0.3.4; Henry \& Wickham, 2020a), \emph{Rcpp} (Version 1.0.6; Eddelbuettel \& François, 2011; Eddelbuettel \& Balamuta, 2018), \emph{readr} (Version 1.4.0; Wickham \& Hester, 2020), \emph{rebus} (Version 0.1.3; Cotton, 2017), \emph{rlang} (Version 0.4.10; Henry \& Wickham, 2020b), \emph{rstan} (Version 2.21.2; Stan Development Team, 2020a), \emph{serCymruTools} (Version 0.1.3; Will, 2021), \emph{StanHeaders} (Version 2.21.0.7; Stan Development Team, 2020b), \emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tibble} (Version 3.1.0; Müller \& Wickham, 2021), \emph{tidyLPA} (Version 1.0.8; Rosenberg et al., 2018), \emph{tidyr} (Version 1.1.3; Wickham, 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al., 2019), and \emph{tsibble} (Version 1.0.0; Wang et al., 2020) for all our analyses, functions are available in the \texttt{serCymruTools} package (Will, 2021), and all other code is available at \url{https://github.com/w-hardy/sercymru}.

@@

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (position_stack).
\end{verbatim}

@@

@@

>>>>

  \begin{itemize}
  \tightlist
  \item
    We used the \texttt{fable} (O'Hara-Wild et al., 2021a), \texttt{fasster} (O'Hara-Wild \& Hyndman, 2021), and the \texttt{fable.prophet} (O'Hara-Wild, 2020) packages to conduct the time series analyses.
  \item
    Time series linear model
  \item
    Decomposition model
  \item
    Seasonal naïve (with and without drift)
  \item
    Autoregressive integrated moving average (ARIMA; Box et al., 2015)
  \item
    Holt-Winters Additive Model (Chatfield, 1978)
  \item
    Prophet (Taylor \& Letham, 2018)
  \item
    Combination models (cf.~Thomson et al., 2019)
  \item
    Prescribing quantities log transformed and forecasts use median values to reduce bias that back transformation would introduce when using the mean
  \item
    Model selection

    \begin{itemize}
    \tightlist
    \item
      Ranked by RMSE and Winkler ({\textbf{???}}), then the best mean ranking model is chosen

      \begin{itemize}
      \tightlist
      \item
        Winkler score penalises models that have observations falling outside the prediction interval, thus rewards models with narrow intervals
      \end{itemize}
    \item
      This combination accounts for both the point and distributional accuracy of the forecasts
    \end{itemize}
  \end{itemize}
\end{itemize}

@@

We used R (Version 4.0.2; R Core Team, 2020) and the R-packages \emph{dplyr} (Version 1.0.4; Wickham, François, et al., 2021), \emph{fable} (Version 0.3.0; O'Hara-Wild et al., 2021a, 2021b; O'Hara-Wild, 2020), \emph{fable.prophet} (Version 0.1.0; O'Hara-Wild, 2020), \emph{fabletools} (Version 0.3.0; O'Hara-Wild et al., 2021b), \emph{fasster} (Version 0.1.0.9100; O'Hara-Wild \& Hyndman, 2021), \emph{feasts} (Version 0.1.7; O'Hara-Wild et al., 2021c), \emph{forcats} (Version 0.5.1; Wickham, 2021), \emph{furrr} (Version 0.2.2; Vaughan \& Dancho, 2021), \emph{future} (Version 1.21.0; Bengtsson, 2020), \emph{ggplot2} (Version 3.3.3; Wickham, 2016), \emph{lubridate} (Version 1.7.10; Grolemund \& Wickham, 2011), \emph{MplusAutomation} (Version 0.8; Hallquist \& Wiley, 2018), \emph{papaja} (Version 0.1.0.9942; Aust \& Barth, 2020), \emph{prophet} (Version 0.6.1; O'Hara-Wild, 2020; Taylor \& Letham, 2020), \emph{purrr} (Version 0.3.4; Henry \& Wickham, 2020a), \emph{Rcpp} (Version 1.0.6; Eddelbuettel \& François, 2011; Eddelbuettel \& Balamuta, 2018), \emph{readr} (Version 1.4.0; Wickham \& Hester, 2020), \emph{rebus} (Version 0.1.3; Cotton, 2017), \emph{rlang} (Version 0.4.10; Henry \& Wickham, 2020b), \emph{rstan} (Version 2.21.2; Stan Development Team, 2020a), \emph{serCymruTools} (Version 0.1.3; Will, 2021), \emph{StanHeaders} (Version 2.21.0.7; Stan Development Team, 2020b), \emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tibble} (Version 3.1.0; Müller \& Wickham, 2021), \emph{tidyLPA} (Version 1.0.8; Rosenberg et al., 2018), \emph{tidyr} (Version 1.1.3; Wickham, 2020), \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al., 2019), and \emph{tsibble} (Version 1.0.0; Wang et al., 2020) for all our analyses, functions are available in the \texttt{serCymruTools} package (Will, 2021), and all other code is available at \url{https://github.com/w-hardy/sercymru}.

@@

\includegraphics{paper_files/figure-latex/all-prescribing-modelling-1.pdf}

@@

\begin{verbatim}
## Warning: Ignoring unknown aesthetics: fill

## Warning: Ignoring unknown aesthetics: fill

## Warning: Ignoring unknown aesthetics: fill
\end{verbatim}

@@

\leavevmode\hypertarget{ref-R-fasster}{}%
O'Hara-Wild, M., \& Hyndman, R. (2021). \emph{Fasster: Fast additive switching of seasonality, trend and exogenous regressors}. \url{https://github.com/mitchelloharawild/fasster}
